---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DASShapefiles

This package contains shapefiles from: <br>  

- [Natura2000](https://www.eea.europa.eu/data-and-maps/data/natura-13/natura-2000-spatial-data/natura-2000-shapefile-1)<br> 
- [Country shape](https://www.diva-gis.org/datadown)<br> 
- [Economic Exclusive zone](https://marineregions.org/gazetteer.php?p=details&id=5669)<br> 
- [ICES rectangles](https://gis.ices.dk/shapefiles/ICES_rectangles.zip)<br> 

The goal is provide easy access to shapefiles on the North and Baltic Sea. 

Shapefiles are in CRS 3035, and 4326.

# 1. Installation

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/DASShapefiles")
```

```{r libreria}
library(DASShapefiles)
```

Other useful packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(sf)
library(ggplot2)
library(ggspatial)
```

# 2. Map

## 2.1. Europe

```{r, fig.width=10}
ggplot2::ggplot()+ ggplot2::geom_sf(data = Europe_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## 2.2. DE_land

To download go to: [DIVA-GIS](https://www.diva-gis.org/datadown)<br> 
Select the country and Administrative areas. 

```{r fig.width=10}
ggplot2::ggplot()+ ggplot2::geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```


## 2.3. DE_EEZ

To download: go to [EMODnet](https://www.emodnet-humanactivities.eu/view-data.php)>Select wind farms in the map.
Select Other Forms of Area Management/Designation>Click on the download icon on [Exclusive Economic Zone](https://www.emodnet-humanactivities.eu/search-results.php?dataname=Exclusive+Economic+Zone).
Select Download Data, Select the country, and the Sector, and the Link should be ready to download.

```{r, fig.width=10}
ggplot2::ggplot()+ ggplot2::geom_sf(data = DE_EEZ,
                                    colour = "red", fill= NA,alpha=0.9, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```


## 2.4. DE_coast

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_coast, 
                   colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 1,linetype="dashed")+
  ggplot2::theme_bw()
```

## 2.5. EEZ North

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_North, 
                                    colour = "black", fill= '#bde0fe',alpha=0.9, lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

## 2.6. DE Baltic 

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_Baltic, 
                                    colour = "black", fill= '#bde0fe',alpha=0.9, lwd = 0.5)+
  ggplot2::theme_bw()
```

# 3. Protected areas

### 3.1. Natura Areas

To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **NatureConservation**

Other option is [eea](https://www.eea.europa.eu/data-and-maps/data/natura-13)

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::theme_bw()
```

### 3.2. Special Conservation Areas

Other option is [Europen Environment Agency](https://www.eea.europa.eu/data-and-maps/data/natura-14)
This file contains a more complete data from the natura 2000 sites.

For using the SCA (Special Conservation Areas) in Germany

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_SCA, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 3.3. SPA and SCA

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::geom_sf(data = DE_SCA, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::theme_bw()
```


# 4. Human activities

## 4.1. Wind Farms

### 4.1.1. BfN
To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **'OffshoreWindEnergy'**

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_OWF, colour = "#fe0001", fill= '#bde0fe',alpha=0.9, size=1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 4.1.2. EMODnet

To download: Go to [EMBO](https://www.emodnet-humanactivities.eu/view-data.php). Select [human activities](https://www.emodnet-humanactivities.eu/search-results.php?dataname=Wind+Farms+%28Polygons%29)

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = OWF_EMODnet, colour = "#fe0001", fill= '#bde0fe',alpha=0.9, size=1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 4.1.3. Map with annotations

```{r}
ggplot(OWF_EMODnet)+ 
  geom_sf(data = DE_EEZ,colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DE_coast,colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = DE_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = OWF_EMODnet, aes(fill= STATUS), colour = "grey",alpha=0.9, size=0.5)+
  scale_fill_manual(values = c("yellow", "blue", "green"))+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme(legend.position = 'bottom')
```


## 4.2. Shipping lines

To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **Shipping**

```{r}
DE_shipping<-DASShapefiles::DE_shipping
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = DE_shipping, colour = "blue", fill= 'blue',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 4.2.1. Map with annotations

```{r}
ggplot(DE_shipping)+ 
  geom_sf(data = DE_EEZ,colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DE_coast,colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = DE_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = DE_shipping, aes(fill= Type), colour = "grey",alpha=0.9, size=0.5)+
  scale_fill_manual(values = c("blue", "deepskyblue", "turquoise"))+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme(legend.position = 'bottom')+
  ggplot2::theme_bw()
```


# 5. Environmental variables

## 5.1. Bathymetry

To download [GEBCO: General Bathymetry Chart of the Oceans](https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area)

Go to: [Download](https://download.gebco.net/) data for user-defined areas
Use the application.
Add your coordinates here I use 1 to 10 and 50 to 60.
Add to basket and download.

For steps to manipulate the data check my [blogpost](https://www.miriam-lerma.com/posts/2025-01-15-bathymetry/)

Reference: If the data sets are used in a presentation or publication then we ask that you acknowledge the source.This should be of the form: GEBCO Compilation Group (2024) GEBCO 2024 Grid (doi:10.5285/1c44ce99-0a0d-5f4f-e063-7086abc0ea0f)

```{r}
Bath_dataframe<-DASShapefiles::DE_bath
```

```{r,message=FALSE}
Bath_dataframe_sub <-Bath_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)%>%
  rename(Bathymetry=3) %>%
  filter(Bathymetry < 10)%>%
  mutate(Bathymetry = as.numeric(Bathymetry))
```

```{r}
DE_land<-st_transform(DE_land, 4326)
```

```{r}
library(ggplot2)
ggplot() +
  geom_raster(data = Bath_dataframe_sub , aes(x = x, y = y, fill = Bathymetry)) +
  geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "mako")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```

Reference: If the data sets are used in a presentation or publication then we ask that you acknowledge the source.This should be of the form: GEBCO Compilation Group (2024) GEBCO 2024 Grid (doi:10.5285/1c44ce99-0a0d-5f4f-e063-7086abc0ea0f)

## 5.2. Distance to coast

Source: [OceanColor NASA](https://oceancolor.gsfc.nasa.gov/resources/docs/distfromcoast/)
I recommend to download **the interpolated 0.01-degree GeoTiff packed together with a brief description file.**
For steps to manipulate the data check my [blogpost](https://www.miriam-lerma.com/posts/2025-02-15-distancetocoast/)

```{r}
DistCoast_dataframe<-DASShapefiles::DE_distancecoast
```

```{r}
library(sf)
```

```{r}
DE_land<-st_transform(DASShapefiles::DE_land, 4326)
```

```{r}
DistCoast_dataframe_sub<-DistCoast_dataframe %>%
  filter(Dist > -20)
```

```{r}
ggplot() +
  geom_raster(data = DistCoast_dataframe_sub, aes(x = x, y = y, fill = Dist)) +
  geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "rocket")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



# 6. Zonification

## 6.1. Study Area

Based on the occurrence of marine animals, the sea areas of the German North Sea and Baltic Sea were divided into study areas, which were then surveyed using specific transect designs.

Source: [BfN](https://nokis.mdi-de-dienste.org/trefferanzeige?docuuid=57fbad61-2960-4fe1-bd28-291894d02c94)

Reference: Dieser Dienst und die darin verfügbaren Daten können gemäß der 'Nutzungsbestimmungen für die Bereitstellung von Geodaten des Bundes' (http://www.gesetze-im-internet.de/bundesrecht/geonutzv/gesamt.pdf) genutzt werden. Quelle: Bundesamt für Naturschutz (BfN) Jahr.


```{r, eval=TRUE}
DE_areas<-DASShapefiles::DE_areas
```

```{r, eval=TRUE}
library(ggplot2)
DE_land<-st_transform(DASShapefiles::DE_land, 4326)
DE_EEZ<-st_transform(DASShapefiles::DE_EEZ, 4326)
DE_coast<-st_transform(DASShapefiles::DE_coast, 4326)
ggplot() +  
   geom_sf(data = DE_EEZ, colour = "black", fill= '#caf0f8', lwd = 0.5)+
  
  geom_sf(data = DE_areas, color='black',fill = '#caf0f8', lwd = 0.5)+
  
  geom_sf(data = DE_areas, aes(fill=stratum),colour = 'transparent')+
  scale_fill_manual(values = c(A = '#7400b8',B = '#6930c3',C = '#5e60ce',D = '#277da1',E = '#577590',F = '#4d908e',
                               G = '#43aa8b',H = '#90be6d',I = '#f9c74f',J = '#ff6d00',K = '#f8961e',L = '#f3722c',
                               N = '#f94144'))+
  
  # Fill colors
 
 
  # Line colors
  geom_sf(data = DE_EEZ, colour = "black", fill= NA, lwd = 0.5)+
  geom_sf(data = DE_coast, colour = "black", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DE_land, colour = '#d9d9d9', fill = '#d9d9d9')+
  geom_sf(data = DE_coast, colour = "black", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  
  coord_sf(xlim = c(3, 15),ylim = c(53, 56))+
  scale_x_continuous(breaks = c(5,7,9,11,13,15),labels = function(x) paste0(x, '\u00B0', "W")) +
  scale_y_continuous(breaks = c(53.5,54.5,55.5,56.5),labels = function(x) paste0(x, '\u00B0', "N"))+
  
  theme(
  legend.position='none',
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "transparent"))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1.5))+
  
  theme(axis.text.x = element_text(size=10,vjust = 15),
        axis.text.y = element_text(size=10,margin = margin(0,-1.50,0,1, unit = 'cm')),
        axis.title = element_blank(),
        axis.ticks.length=unit(-0.25, "cm"))+
  annotate(geom = "text", x = 4.5, y = 55.5, label = "A",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 5.5, y = 55.3, label = "B",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 6.8, y = 55.0, label = "C",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 7.9, y = 54.8, label = "D",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 8.1, y = 54.1, label = "E",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 7.1, y = 54.1, label = "F",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 6.4, y = 54.5, label = "G",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 5.5, y = 54.8, label = "H",size = 4,fontface = 'bold',color='#343a40')+  
  annotate(geom = "text", x = 10.3,y = 54.8, label = "I",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 11,  y = 54.6, label = "J",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 12,  y = 54.5, label = "K",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 12.8,y = 54.8, label = "L",size = 4,fontface = 'bold',color='#343a40')+
  annotate(geom = "text", x = 14.2,y = 54.3, label = "O",size = 4,fontface = 'bold',color='#343a40')
```


## 6.2. ICES Rectangles

Source: [ICES rectangles](https://gis.ices.dk/sf/index.html)<br> 
To download: go to link> click on **Quick Downloads**> select ICES Statistical Rectangles.

```{r}
DE_ICES<-DASShapefiles::DE_ICES
```

```{r}
ggplot2::ggplot(DE_ICES)+ 
  ggplot2::geom_sf(data = DE_ICES, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::geom_sf_text(ggplot2::aes(label =ICESNAME),size=3,family="sans")+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## 6.3. Grids

This grids were created in R and are made available by this package.

To cite: Lerma, M (2023) German North Sea package. 

```{r}
library(ggplot2)
library(ggspatial)
```

### 6.3.1. Grid 5x5

```{r}
ggplot2::ggplot()+ 
  geom_sf(data = DASShapefiles::DE_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DASShapefiles::DE_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DASShapefiles::DE_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::geom_sf(data=DE_grid5x5_3035, colour = "black", fill= NA, lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 6.3.2. Grid 10x10

```{r, eval=FALSE, echo=FALSE}
# Packages
library(tidyverse)
library(sf)

# Create limits
my_data<-data.frame(Longitude=c(3790000,4250000),
Latitude=c(3350000,3680000)
)

# Tranform to CRS
my_points<-my_data %>%
  st_as_sf(coords=c('Longitude','Latitude'),
           #crs=4326,
           crs=3035,
           remove=FALSE)

my_grid<-st_make_grid(my_points, 
                       c(10000, 10000), 
                       what = "polygons",
                       square = TRUE)
# Add id per grid
my_grid_sf = st_sf(my_grid) %>%
  dplyr::mutate(grid_id = 1:length(lengths(my_grid)))

# Check grid
ggplot()+
  geom_sf(data=my_grid, colour = "black", fill= '#56C1D5', lwd = 0.5)

# Asign CRS
grid10x10_3035<-my_grid_sf
```

```{r}
ggplot() +
  geom_sf(data = DASShapefiles::DE_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DASShapefiles::DE_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DASShapefiles::DE_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::geom_sf(data=DE_grid10x10_3035, colour = "black", fill= '#56C1D5', lwd = 0.5, alpha=0.1)+
    xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

### 6.3.3. Grid 10x10 Only EEZ

```{r, eval=FALSE, echo=FALSE}
# Package
library(DASShapefiles)
library(tidyverse)
library(sf)
# Shapefiles
Grid_10km<-grid10x10_3035
Grid_10km_EEZ <- sf::st_intersection(Grid_10km, German_EEZ)
## With data
## Create data frame of grids with data by dropping geometry
Grid_names<-Grid_10km_EEZ %>% 
  st_drop_geometry()
## Select column with grid_id
Grid_names<-Grid_names %>%
  select(grid_id,OBJECTID)
## Grid inside EEZ
Grid_10km_keep<-merge(Grid_10km,Grid_names,by='grid_id',all=TRUE)
## Keep only with data
Grid_10km_keep<- Grid_10km_keep[stats::complete.cases(Grid_10km_keep$OBJECTID),]
# Asign name
grid10x10_EEZ<-Grid_10km_keep
# rdata
save(grid10x10_EEZ, file = "grid10x10_EEZ.RData", version = 2)

##devtools::document()
##devtools::check()
```

```{r}
ggplot2::ggplot()+ 
  geom_sf(data = DASShapefiles::DE_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DASShapefiles::DE_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DASShapefiles::DE_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::geom_sf(data=DE_grid10x10_EEZ, colour = "black", fill= NA, lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  ggplot2::theme_bw()
```

# 7. Different CRS


## 7.1. CRS 3050

```{r}
ggplot() +
  geom_sf(data = DASShapefiles::DE_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DASShapefiles::DE_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DASShapefiles::DE_land, colour = 'black', fill = '#ffffbe')+

  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



## 7.2. CRS 4326

For using other CRS, you can use the function **st_transform** from the package **sf**.

```{r}
library(sf)
```

```{r}
library(DASShapefiles)
library(ggspatial)
```

```{r}
DE_EEZ<-st_transform(DE_EEZ, 4326)
DE_coast<-st_transform(DE_coast, 4326)
DE_land<-st_transform(DE_land, 4326)
DE_natura<-st_transform(DE_natura, 4326)
```

```{r, fig.height=8, fig.width=8}
ggplot()+
  geom_sf(data = DE_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = DE_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = DE_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = DE_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+

  ggspatial::annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

# 8. Resources

- [Marine Spatial Planning](https://www.bsh.de/EN/TOPICS/Offshore/Maritime_spatial_planning/maritime_spatial_planning_node.html)

- [Shipping intensity](https://www.emodnet-humanactivities.eu/download-data.php)
- [Marine Traffic](https://www.marinetraffic.com/en/ais/home/centerx:8.6/centery:53.9/zoom:9)

- [Offshore wind farms](https://map.4coffshore.com/offshorewind/)
- [Windkraftanlagen](https://commons.wikimedia.org/wiki/File:Karte_Offshore-Windkraftanlagen_in_der_Deutschen_Bucht.png)

- [GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)
- [NOAA-GRIDDAP-GLOBAL](https://erddap.ifremer.fr/erddap/griddap/index.html?page=1&itemsPerPage=1000)
- [COPERNICUS](https://resources.marine.copernicus.eu/products)
